\emph{Gaia} DR2 identifies VFTS 682 as the most massive runaway star known to date

### 2018-05-14 VFTS682.tex
minor

\section{Validation with HST proper motions}

Recently, \citet{platais:18} used HST observations (based on GO-12499
\& 13359 observations with WFC3/UVIS and on archival WFPC2 and
GO-12499 with both cameras, \citealt{platais:15}) to determine the
proper motion of massive stars in the 30 Doradus region. For
large distances ($\gtrsim50$\,kpc), HST data can compete or even
exceed Gaia data in quality.

In the catalog from \citet{platais:18} they listed the proper motion
components of VFTS682 relative to the field, providing a
measure of the motion of this star completely independent from
ours. HST images of VFTS682 are not saturated in any epoch \todo{?},
thanks to the large extinction toward this star. They found proper
motion components relative to the field
$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
  yr^{-1}}$ and
$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
  yr^{-1}}$.

@@

\todo{redo below with average HST+Gaia?Actually redo also calculation
  of $v_\mathrm{pec}$...}
\subsection{Does it come from the R136 cluster?}
\label{sec:r136_origin}

@@

Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
based on the kinematic properties of VFTS 682, we would expect several
stars with initial masses larger than $\sim$$150\,M_\odot$ in the
cluster R136.
This is consistent with the detection
of extremely massive stars by \cite{crowther:10} in the core of the
cluster. The projected rotational equatorial
velocity\footnote{However, the determination of the rotational
  velocity for stars showing lines in emission is complicated because
  of the optically thick wind screening the surface of the star, and should be
  considered with caution.} of VFTS 682
reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
line with the average rotation rate of massive stars in the region
\citep[][]{ramirez-agudelo:15}. This suggests that VFTS 682 (i) has not
experienced rotationally induced chemically homogeneous evolution
\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
data of the VFTS survey rule out the presence of a close companion at the
present day. Moreover, the spectral type of VFTS 682
\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
three most massive stars detected in the core of the cluster by
\cite{crowther:10}, with an astonishing similarity in particular with
the spectrum of R136a3. Therefore,
VFTS 682 might be an ideal target to constrain the stellar physics of
stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
it an easier target for observations compared to the similar stars
present in the crowded core of R136.

>>>>

\section{Validation with HST proper motions}
\todo{maybe rename to Astrometry from HST and make sec. 2.1.1 or 2.2
  but above Gaia data?}
Recently, \citet{platais:18} used HST observations (based on GO-12499
\& 13359 observations with WFC3/UVIS and on archival WFPC2 and
GO-12499 with both cameras, see \citealt{platais:15}) to determine the
proper motion of massive stars in the 30 Doradus region. Our star of
interest is not saturated in any HST image \todo{?},
because of the large extinction toward this star. For
large distances ($\gtrsim50$\,kpc), HST data can compete or even
exceed Gaia data in quality.

The catalog from \citet{platais:18} lists the proper motion
components of VFTS682 relative to the field, providing a
measure of the motion of this star completely independent from
ours. They found proper
motion components relative to the field
$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
  yr^{-1}}$ and
$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
  yr^{-1}}$.

@@

\todo{redo below with average HST+Gaia?Actually redo also calculation
  of $v_\mathrm{pec}$}
\subsection{Does it come from the R136 cluster?}
\label{sec:r136_origin}

@@

Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
based on the kinematic properties of VFTS 682, we would expect several
stars with initial masses larger than $\sim$$150\,M_\odot$ in the
cluster R136.
This is consistent with the detection
of extremely massive stars by \cite{crowther:10} in the core of the
cluster. The projected rotational equatorial
velocity\footnote{However, the determination of the rotational
  velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
  considered with caution.} of VFTS 682
reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
line with the average rotation rate of massive stars in the region
\citep[][]{ramirez-agudelo:15}. This suggests that VFTS 682 (i) has not
experienced rotationally induced chemically homogeneous evolution
\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
data of the VFTS survey rule out the presence of a close companion at the
present day. Moreover, the spectral type of VFTS 682
\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
three most massive stars detected in the core of the cluster by
\cite{crowther:10}, with an astonishing similarity in particular with
the spectrum of R136a3. Therefore,
VFTS 682 might be an ideal target to constrain the stellar physics of
stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
it an easier target for observations compared to the similar stars
present in the crowded core of R136.
